---
title: "Metro market assessment"
author: "Sifan"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide

# runtime:shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, include = T)
# rmarkdown::render("Report-template.Rmd",output_file = paste0(target_cbsa,"_",Sys.Date(),'.html'))
```

```{r func, include=FALSE, echo=F}
source("R/pip.R")

quick_bar = function(df, var,HL = "0",wt = F, scale = 1, acc = 0.1){
  df <- df %>% 
    mutate_(value = var) %>%
    mutate(value = value*scale) %>%
    mutate(highlight = (cbsa_code == HL))
  
  if (wt) {df <- df %>% mutate(value = value/cbsa_emp)}
  
  metro.data::bbplot(df,aes(x = reorder(cbsa_name,value), y = value, 
                label = scales::comma(value,acc), fill = highlight))+
  geom_col()+
  geom_text()+
  scale_fill_brewer(palette = "Set3", guide = F)+
  coord_flip()+
  ggtitle(var)
  
  # return(df)
}

```

```{r target, include=T}

# change target metro and peers here
# target_cbsa <- "19740"
# target_cbsa <- "31140"
target_cbsa <- "24340"
# peer_cbsa <- c("35620", "16980", "31080", "47900")
peer_cbsa <- ""
# peer_cbsa <- c("13820","16740","17140","18140","24660","26900","28140","34980", "36540", "32820")
# peer_cbsa <- c("12420", "33460", "38060", "38900", "41740", "41620", "42660")

# get counties
target_co <- (county_cbsa_st %>%
  filter(cbsa_code == target_cbsa) )[["stco_code"]]

peer_co <- (county_cbsa_st %>%
  filter(cbsa_code %in% peer_cbsa) )[["stco_code"]]

# filter datasets
cbsa_output <- cbsa_all %>%
  filter(cbsa_code %in% c(target_cbsa, peer_cbsa)) %>%
  left_join(county_cbsa_st[c("cbsa_code", "cbsa_emp")], by = "cbsa_code") %>%
  unique()

# make it a shiny app!
# selectInput("target_cbsa", "", )

```


## Dynamism


### Shift share (job change by industry above/below national average)--for metro
It's easier to get from EMSI directly: Industry - Inustry Table - Shiftshare. You can toggle between different industry levels

### Tradable share of employment--metro and peers

```{r traded}

emp <- read_csv("../../_metro_data_warehouse/data_raw/CBP EMP/efsy_cbp_2016.csv")

emp %>%
  mutate(stco_code = paste0(str_pad(fipstate, 2, "left", "0"), str_pad(fipscty,3,"left",0))) %>%
  filter(stco_code %in% c(target_co, peer_co)) %>%
  left_join(metro.data::naics_xwalk, by = c("naics" = "code_naics6_all")) %>%
  filter(!is.na(traded.naics6)) %>%
  left_join(county_cbsa_st[c("stco_code", "cbsa_code", "cbsa_name")]) %>%
  group_by(cbsa_code, cbsa_name,traded.naics6 ) %>%
  summarise(emp = sum(emp)) %>%
  pivot_wider(names_from = "traded.naics6", values_from = "emp") %>%
  mutate(pct_traded = Traded/(Traded+Local)) %>%
  knitr::kable()
  


```

### Change in share of employment in Advanced Industries--metro and peers


```{r ai}
emp06 <- read_csv("../../_metro_data_warehouse/data_raw/CBP EMP/efsy_cbp_2006.csv")

pct_ai <- function(df)df %>%
  mutate(stco_code = paste0(str_pad(fipstate, 2, "left", "0"), str_pad(fipscty,3,"left",0))) %>%
  filter(stco_code %in% c(target_co, peer_co)) %>%
  mutate(naics = gsub("\\/", "", naics))%>%
  filter(str_length(naics)== 4)%>%
  left_join(metro.data::naics4_ai, by = c("naics" = "naics4_code")) %>%
  # mutate(naics4_aitype = ifelse(is.na(naics4_aitype), NA, 1))%>%
  left_join(county_cbsa_st[c("stco_code", "cbsa_code", "cbsa_name")]) %>%
  group_by(cbsa_code, cbsa_name,naics4_aitype) %>%
  summarise(emp = sum(emp)) %>%
  # mutate(pct = emp/sum(emp))
  pivot_wider(names_from = "naics4_aitype", values_from = "emp") %>%
  mutate(pct_ai = (aiEnergy + aiMfg + aiServ)/(aiEnergy + aiMfg + aiServ +`NA`)) %>%
  select(cbsa_code, cbsa_name, pct_ai)

pct_ai(emp06)%>%
  rename(pct_ai_06 = pct_ai)%>%
  left_join(pct_ai(emp), by = c("cbsa_code", "cbsa_name")) %>%
  rename(pct_ai_16 = pct_ai) %>%
  knitr::kable()

```


### R&D per 1,000 workers--metro and peers
Annual average R&D expenditure at higher education institutions per 1000 workers, 2012 - 2016
(plots can be generated for other tables as well, if needed)
```{r RD}

cbsa_output %>%
  select(cbsa_code, cbsa_name, rd_total, cbsa_emp) %>%
  mutate(rd_per_1000worker = rd_total/cbsa_emp * 1000/5) %>%
  knitr::kable()

quick_bar(cbsa_output, "rd_total",HL = target_cbsa, wt = T, scale = 200)

```


### Do we have rankings for UofL by research focus similar to UAB stats?

Annual R&D expenditure at UofL, thousand US dollar
```{r RD univ}
load("univ_rd_detail/univ_rd_18.rda")

univ_rd_18 %>%
  filter(stco_code %in% target_co) %>%
  select(year, univ_name, field_broad, field_detail, total_rd)%>%
  unique()%>%
  filter(field_detail == "Total") %>%
  pivot_wider(names_from = "year", values_from = "total_rd") %>%
  DT::datatable()

```

### Startups launched at universities--metro and peers
Number of start-ups launched at universities, 2006 - 2017
```{r startups}
cbsa_output %>%
  select(cbsa_code, cbsa_name, tot_st) %>%
  knitr::kable()

```

### INC 5000 high growth companies--metro and peers
Number of Inc.5000 high growth companies per 1M workers, 2011 - 2017

```{r inc5000}
cbsa_output %>%
  select(cbsa_code, cbsa_name, i5hgc_density) %>%
  knitr::kable()
```


### Startups per 1,000 workers--metro and peers
```{r startup}
load("../crunchbase/cb-result/data/cb_cbsa_index.rda")

index %>%
  filter(cbsa_code %in% c(target_cbsa, peer_cbsa))%>%
  select(cbsa_code, cbsa_name, msa_total, SCI = div) %>%
  unique()%>%
  left_join(cbsa_output[c("cbsa_code", "cbsa_emp")], by = "cbsa_code") %>%
  mutate(startup_per_million = msa_total/cbsa_emp*1000000)%>%
  knitr::kable()

```

## Skills
### % of population 25+ with AA/BA--metro and peers
```{r edu}
cbsa_output %>%
  select(cbsa_code, cbsa_name, pct_edu_associate, pct_edu_baplus)%>%
  knitr::kable()
  
  
```

### Employment-to-population ratio 18-64--metro and peers

```{r epratio}

cbsa_output %>% 
  select(cbsa_code, cbsa_name, epratio_total)%>%
  knitr::kable()


```


### Prevalence of health problems--metro and peers
```{r health}
library(RSocrata)
token <- Sys.getenv("RSocrata_token")

cities <- paste0(gsub("\\,.+","",cbsa_output$cbsa_name_short), collapse = "','")

cityhealth <- read.socrata(paste0("https://chronicdata.cdc.gov/resource/csmm-fdhi.csv?category=Health Outcomes&$where=cityname in",
                                  "('",cities,"')"),token)

# peer comparison
cityhealth %>% 
  filter(measureid %in% c("MHLTH", "PHLTH")) %>%
  filter(geographiclevel == "City") %>%
  filter(datavaluetypeid == "AgeAdjPrv") %>%
  select(cityname, measure, data_value, year) %>%
  group_by(cityname, measure)%>%
  summarise(data_value = mean(data_value))%>%
  pivot_wider(names_from = "measure", values_from = "data_value") %>%
  knitr::kable()
```

### Struggling families? (I think this may not be available?)--metro and peers

not available

### Share of workers who are low-wage--metro and peers

```{r low-wage}
cbsa_low_wage_worker %>%
  filter(cbsa_code %in% c(target_cbsa, peer_cbsa)) %>%
  filter(population %in% c("Low-wage workers","Workers"))%>%
  select(cbsa_code, cbsa_name, population, total)%>%
  pivot_wider(names_from = "population", values_from = "total") %>%
  mutate(pct_low_wage = `Low-wage workers`/`Workers`) %>%
  knitr::kable()
```


### Share of population out-of-work--metro and peers
Note: metro aggregates only include core counties, as listed
```{r oow}

names <- co_oow %>%
  filter(cbsa_code %in% c(target_cbsa, peer_cbsa)) %>%
  select(cbsa_code, cbsa_name, pl_name) %>%
  unique()%>%
  group_by(cbsa_code) %>%
  summarise(cbsa_name = last(cbsa_name),
            core_counties = paste0(pl_name, collapse = ", "))


co_oow %>%
  filter(cbsa_code %in% c(target_cbsa, peer_cbsa))%>%
  filter(population %in% c("Out-of-work population","Universe population")) %>%
  select(stco_code, age, cbsa_code, cbsa_name, pl_name, population, total) %>%
  group_by(cbsa_code, population, age) %>%
  summarise(total = sum(total))%>%
  pivot_wider(names_from = "population", values_from = "total") %>%
  mutate(pct_oow = `Out-of-work population`/`Universe population`) %>%
  select(-contains("population")) %>%
  pivot_wider(names_from = "age", values_from = "pct_oow")%>%
  left_join(names, by = "cbsa_code")%>%
  knitr::kable()
```


### Digital skills demand, 2002/2016--metro


```{r digital, echo=FALSE}
cbsa_output %>% 
  select(cbsa_code, cbsa_name,contains("cbsa_pct"), -cbsa_pct_exports_gdp)%>%
  knitr::kable()

```

## Access
### Job density change by block group?--metro
we don't have block group data in the warehouse yet - can ask Joanne

### Geographic of prosperity trend (UVA data)--metro + MW/Southeastern trendlines

http://statchatva.org/changing-shape-of-american-cities/

### Prevalence of health problems by census tract?--metro


```{r map, include=FALSE}
library(sf)
library(tmap)
tmap_mode("view")

state <- county_cbsa_st%>%
  filter(cbsa_code == target_cbsa)%>%
  select(st_code)%>%
  unique()%>%
  pull(st_code)

# map <- tidycensus::get_acs(geography = "tract", variables = "B01003_001", state = state, geometry = T)
# test2 <- tidycensus::get_acs(geography = "tract", variables = "B01003_001", state = "18", geometry = T)
# map2 <- rbind_list(list(test, test2))

map <- purrr::map(state, ~ tidycensus::get_acs(geography = "tract", variables = "B01003_001", state = .x, geometry = T))
# 
# length(map)
map <- rbind(map[[1]], map[[2]])

```


Only available for core city
```{r}

tract_health <- cityhealth %>%
  filter(measureid %in% c("MHLTH", "PHLTH")) %>%
  filter(geographiclevel == "Census Tract") %>%
  mutate(stcotr_code = str_pad(tractfips, 11, "left","0"))%>%
  select(year, stcotr_code, measureid, data_value) %>%
  pivot_wider(names_from = "measureid", values_from = "data_value")

map_health <- map %>%
  mutate(stco_code = str_sub(GEOID,1,5)) %>%
  left_join(tract_health, by = c("GEOID" = "stcotr_code"))

tm_shape(map_health %>%
  filter(stco_code %in% target_co) %>%
  filter(!is.na(year)))+
  tm_polygons(c("MHLTH", "PHLTH"), alpha = 0.7)+
  tm_facets(nrow = 1, sync = T)

```

### Number of jobs reachable by travel time (including public transit)--metro
Number of jobs reachable by 30 minutes auto and 30 minutes transit
```{r access}
cbsa_output %>% 
  select(cbsa_code, cbsa_name,auto_30, transit_30, ratio_30)%>%
  knitr::kable()
```


map: http://access.umn.edu/research/america/transit/2018/maps/index.html

### Are the NCES data behind the proficiency map downloaded and generally available, or did you do those special for BHM?
You can download school level data from NCES. we aggregated school level data to various geography. 
```{r school}

tract_school <- final %>%
  mutate(stco_code = str_sub(fips,1,5)) %>%
  filter(stco_code %in% target_co) %>%
  mutate(stcotr_code = str_sub(fips, 1,11)) %>%
  group_by(stcotr_code) %>%
  summarise(pct_pass = sum(ALL_TOTAL_passed_H)/sum(ALL_TOTAL_total))

map_school <- map %>%
  mutate(stco_code = str_sub(GEOID,1,5)) %>%
  left_join(tract_school, by = c("GEOID" = "stcotr_code"))


tm_shape(map_school%>%
  filter(str_sub(GEOID,1,5) %in% target_co) %>%
    mutate(lab = scales::percent(pct_pass)))+
  tm_polygons(col = "pct_pass", id = "lab", alpha = 0.7)

# map_school %>%
#   filter(str_sub(GEOID,1,5) %in% target_co)%>%
#   ggplot(aes(fill = pct_pass))+
#   scale_fill_distiller(type = "div", palette = "PiYG")+
#   geom_sf()+
#   coord_sf(crs = 26911) 


```


